Decoding the Amplitude-Modulated Part the DCF77 Longwave Time Signal by using the Goertzel Algorithm

### 2024-08-22 paper/main.tex
Added comments to the script. Added reference for EMA to the paper

\subsection{The detector}
The output of the Goertzel algorithm is already following the waveform of the binary time signal but still includes fluctuations due to noise and fading.
To reconstruct the original waveform of the binary time signal, it is processed inside a detector.
As the values of the Goertzel algorithm will follow the pattern of the signal's amplitude being switched between \SI{15}{\percent} and \SI{100}{\percent}, the detection can be done by comparing the value of the current sample to a threshold.
To compensate the changing of magnitude of the Goertzel algorithm's values that is caused by the current receiving situation, an Exponential Moving Average (EMA) is calculated by using \eqref{eqn:ema} and employed to create an adaptive threshold.
\begin{equation}
    y[t] = \alpha \cdot x[t] + (1 - \alpha) \cdot y[t-1]
    \label{eqn:ema}
\end{equation}
This part can also be represented as a block diagram:
\begin{figure}[h]
    \centering
    \resizebox{0.4\textwidth}{!}{
        \begin{tikzpicture}[auto, node distance=2cm]
            %Defining a cross set for the multiplication elements
            \tikzset{cross/.style={path picture={ 
            \draw[black]
            (path picture bounding box.south east) -- (path picture bounding box.north west) (path picture bounding box.south west) -- (path picture bounding box.north east);
            }}}

@@

\section{MATLAB Simulation of the Receiver}
To verify and underline the working principle of the receiver, a simulation of the whole system was created using MATLAB.
The whole simulation was divided into five sections to allow for a clear structure:
\begin{enumerate}
    \item Simulation Parameter Section
    \item Signal Creation Section
    \item Goertzel Algorithm Section
    \item Detection Section
    \item FFT and Plot Section
\end{enumerate}

@@

\begin{thebibliography}{00}
\bibitem{b1} -, ``DCF77,'' 2024, Wikipedia entry. [Online]. Available: https://de.wikipedia.org/wiki/DCF77
\bibitem{b2} D. Priester and P. Hetzel and A. Bauch, ``Zeit- und Normalfrequenzverbreitung mit DCF77,'' PTB Mitteilungen, 2004, [Online]. Available: https://www.ptb.de/cms/fileadmin/internet/fachabteilungen/abteilung\_4/4.4\_zeit\_und\_frequenz/4.42/dcf77.pdf
\bibitem{b3} D. Priester and P. Hetzel and A. Bauch, ``Zeit- und Normalfrequenzverbreitung mit DCF77: 1959-2009 und darüber hinaus,'' PTB Mitteilungen, 2009, [Online]. Available: https://www.ptb.de/cms/fileadmin/internet/fachabteilungen/abteilung\_4/4.4\_zeit\_und\_frequenz/pdf/2009\_Bauch\_PTBM\_\_DCF77.pdf
\bibitem{b4} P. Hetzel, ``Time dissemination via the LF transmitter DCF77 using a pseudo-random phase-shift keying of the carrier,'' PTB Braunschweig, 1988, [Online]. Available: https://www.ptb.de/cms/fileadmin/internet/fachabteilungen/abteilung\_4/4.4\_zeit\_und\_frequenz/pdf/5\_1988\_Hetzel\_-\_Proc\_EFTF\_88.pdf
\bibitem{b5} -, ``DCF77 webpage,'' 2024, Official page of the PTB [Online]. Available: https://www.ptb.de/cms/en/ptb/fachabteilungen/abt4/fb-44/ag-442/dissemination-of-legal-time/dcf77.html
\bibitem{b6} R. G. Lyons, ``Understanding Digital Signal Processing,'' Third Edition, Prentice Hall, 2010.
\bibitem{b9} J. Hoffmann, ``Sampling bandpass signals,'' 2021, Internet Article. [Online]. Available: https://www.dsprelated.com/showarticle/1407.php
\bibitem{b10} F. Zaplata and M. Kasal, ``SDR Implementation for DCF77,'', 2013, IEEE Paper. [Online]. Available: https://ieeexplore.ieee.org/document/6530943
\bibitem{b7} -, ``Goertzel algorithm,'' 2024, Wikipedia entry. [Online]. Available: https://en.wikipedia.org/wiki/Goertzel\_algorithm
\bibitem{b8} K. Banks, ``The Goertzel Algorithm,'' 2002, Internet Article. [Online]. Available: https://courses.cs.washington.edu/courses/cse466/12au/calendar/Goertzel-EETimes.pdf
\end{thebibliography}
\end{document}

>>>>

\subsection{The detector}
The output of the Goertzel algorithm is already following the waveform of the binary time signal but still includes fluctuations due to noise and fading.
To reconstruct the original waveform of the binary time signal, it is processed inside a detector.
As the values of the Goertzel algorithm will follow the pattern of the signal's amplitude being switched between \SI{15}{\percent} and \SI{100}{\percent}, the detection can be done by comparing the value of the current sample to a threshold.
To compensate the changing of magnitude of the Goertzel algorithm's values that is caused by the current receiving situation, an Exponential Moving Average (EMA) is calculated by using \eqref{eqn:ema} and employed to create an adaptive threshold \cite{b11}.
\begin{equation}
    y[t] = \alpha \cdot x[t] + (1 - \alpha) \cdot y[t-1]
    \label{eqn:ema}
\end{equation}
This part can also be represented as a block diagram:
\begin{figure}[h]
    \centering
    \resizebox{0.4\textwidth}{!}{
        \begin{tikzpicture}[auto, node distance=2cm]
            %Defining a cross set for the multiplication elements
            \tikzset{cross/.style={path picture={ 
            \draw[black]
            (path picture bounding box.south east) -- (path picture bounding box.north west) (path picture bounding box.south west) -- (path picture bounding box.north east);
            }}}

@@

\section{MATLAB Simulation of the Receiver}
To verify and underline the working principle of the receiver, a simulation of the whole system was created using MATLAB.
The whole simulation was divided into five sections to allow for a clear structure:
\begin{enumerate}
    \item Simulation Parameter Section
    \item Signal Creation Section
    \item Goertzel Algorithm Section
    \item Detection Section
    \item FFT and Plot Section
\end{enumerate}
These sections will be discussed in the following.

\subsection{Simulation Parameter Section}
In the Simulation Parameter Section,

@@

\begin{thebibliography}{00}
\bibitem{b1} -, ``DCF77,'' 2024, Wikipedia entry. [Online]. Available: https://de.wikipedia.org/wiki/DCF77
\bibitem{b2} D. Priester and P. Hetzel and A. Bauch, ``Zeit- und Normalfrequenzverbreitung mit DCF77,'' PTB Mitteilungen, 2004, [Online]. Available: https://www.ptb.de/cms/fileadmin/internet/fachabteilungen/abteilung\_4/4.4\_zeit\_und\_frequenz/4.42/dcf77.pdf
\bibitem{b3} D. Priester and P. Hetzel and A. Bauch, ``Zeit- und Normalfrequenzverbreitung mit DCF77: 1959-2009 und darüber hinaus,'' PTB Mitteilungen, 2009, [Online]. Available: https://www.ptb.de/cms/fileadmin/internet/fachabteilungen/abteilung\_4/4.4\_zeit\_und\_frequenz/pdf/2009\_Bauch\_PTBM\_\_DCF77.pdf
\bibitem{b4} P. Hetzel, ``Time dissemination via the LF transmitter DCF77 using a pseudo-random phase-shift keying of the carrier,'' PTB Braunschweig, 1988, [Online]. Available: https://www.ptb.de/cms/fileadmin/internet/fachabteilungen/abteilung\_4/4.4\_zeit\_und\_frequenz/pdf/5\_1988\_Hetzel\_-\_Proc\_EFTF\_88.pdf
\bibitem{b5} -, ``DCF77 webpage,'' 2024, Official page of the PTB [Online]. Available: https://www.ptb.de/cms/en/ptb/fachabteilungen/abt4/fb-44/ag-442/dissemination-of-legal-time/dcf77.html
\bibitem{b6} R. G. Lyons, ``Understanding Digital Signal Processing,'' Third Edition, Prentice Hall, 2010.
\bibitem{b9} J. Hoffmann, ``Sampling bandpass signals,'' 2021, Internet Article. [Online]. Available: https://www.dsprelated.com/showarticle/1407.php
\bibitem{b10} F. Zaplata and M. Kasal, ``SDR Implementation for DCF77,'', 2013, IEEE Paper. [Online]. Available: https://ieeexplore.ieee.org/document/6530943
\bibitem{b7} -, ``Goertzel algorithm,'' 2024, Wikipedia entry. [Online]. Available: https://en.wikipedia.org/wiki/Goertzel\_algorithm
\bibitem{b8} K. Banks, ``The Goertzel Algorithm,'' 2002, Internet Article. [Online]. Available: https://courses.cs.washington.edu/courses/cse466/12au/calendar/Goertzel-EETimes.pdf
\bibitem{b11} R. G. Lyons, ``Setting the 3-dB Cutoff Frequency of an Exponential Averager,'' 2012, Internet Article. [Online]. Available: https://www.dsprelated.com/showarticle/182.php
\end{thebibliography}
\end{document}
